---
title: Figures, Tables and Crossref
format:
    html:
        toc: true
--- 


## Figures

Different ways to deal with figures depending on the format and storage. It can be:

- Internal Figure: Using code chunks to generate graphics e.g., using ggplot2, pmplots, etc. 
- External Figure: Sourced from png, pdf, etc. 

For more info <https://quarto.org/docs/authoring/figures.html> 

### Internal Figures 

```{r}
#| echo: fenced
#| eval: true
#| label: fig-pk
#| fig-width: 10
#| fig-cap: this is the caption
#| output-location: column-fragment

library(ggplot2)

pk.profile <- ggplot(Theoph, aes(x=Time, y=conc)) + 
    geom_point(color="blue", 
               alpha=0.6)

pk.profile
```

The `#| label` option is how the figure is identified when you cross reference it. To cross reference the figure in text type `@label`

-   @fig-pk is my first Figure 

-   The "fig-cap" is the figure's title

-   To display the figure code and output figure side-by-side use the option: "output-location: column-fragment"

```{r}
#| echo: fenced
#| eval: true
#| label: fig-pk
#| fig-width: 5
#| fig-cap: this is the caption
#| output-location: column-fragment

library(ggplot2)

PK_profile<-ggplot(Theoph, aes(x=Time, y=conc)) + 
    geom_point(color="blue", 
               alpha=0.6)

PK_profile